---
title: "SDS 3386"
subtitle: "Assignment 4"
author: "Thinh Hong"
date: "`r Sys.time()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---


```{r, warning=FALSE, message=FALSE}
# load libraries
library(dplyr)
library(janitor)
library(ggplot2)
library(skimr)
library(tidyverse)
library(broom)
library(readxl)
```
# Description
 By studying the data in the statementofvotescastoctober242018.xls,it is being deterrmined if that as the number of people per household increases, the relative amount of votes for the winner decreases. The exact relationship between them may differ by area. The amount of votes the winner had in every ward is compared to the total amount of votes in that ward. Different area typs are compared 
# Headline

The following plots is the household density vs the fraction of the winner votes over the total votes. All the Ward winners were split into three sections by the householddensity. The first plot contains all ward winners information with 0-(1/3) of the full range in terms of household density. First plot is called urban. Second plot is similar except it is split from 1/3 - 2/3 and is called sub-urban.
Lastly the last plot is split from 2/3 to 3/3 and is called rural.

```{r, warning=FALSE, message=FALSE}


dataset <- read_excel("statementofvotescastoctober242018.xls", sheet = "Summary",skip = 4)

# Cleaning dataset and getting the turnout rate
dataset <- dataset[-c(1,2),]
dataset <- dataset %>%
  select(-(...3))%>%
  filter(!grepl('Adv',Precinct))%>%
  clean_names() %>%
  mutate(ward = as.numeric(str_sub(precinct, 0, 2)))%>%
  group_by(ward) %>%
  summarise(
    registered_voters = sum(registered_voters),
    cards_cast = sum(cards_cast)
  ) %>%                                                  
  mutate(fraction_voted = cards_cast / registered_voters) %>%
  na.omit()

wardCleaner <- function(sheetNumber){
  # reads datasheet
  data <- read_excel("statementofvotescastoctober242018.xls", sheet = sheetNumber,skip = 2)
  #Cleans data
  data <- data %>% 
    clean_names() %>%
    select(-starts_with("x")) %>%
    select(-precinct_7,-overvotes,-undervotes,-times_cast) %>%
    filter(precinct_1 != "City / Ville") %>%
    filter(precinct_1 != "Ottawa") %>%
    filter(!grepl('Adv',precinct_1))%>%
    filter(!str_detect(precinct_1, "City / Ville - Total")) %>%
    rename(precinct = precinct_1)
  return (data)
}

# Function to get winner and runnerUp votes
GetWinnerResults <- function(ward) {
  # gets the cleaned sheet data
  sheet_data <- wardCleaner(str_c('Ward ', toString(ward)))
  winner_results <- sheet_data %>%
    select(-precinct) %>%
    select(-total_votes) %>%
    summarise_each(sum) %>%
    # addts new colums
    pivot_longer(
      cols = everything(), 
      names_to = 'candidate',
      values_to = 'votes'
    ) %>% 
    ## summarise and gets total votes for winner and runner up
    summarise(
      winner = candidate[which.max(votes)],
      total_votes = sum(votes),
      votes_winner = max(votes)
    ) %>%
    # adds fractions of votes for winner, runner up and how much the runner up won by
    mutate(
      fraction_of_votes = votes_winner / total_votes,
      ward = ward
      )  
  # Selects important columns for this hypothesis
  winner_results <- winner_results %>% 
    select(ward, winner,votes_winner,total_votes,fraction_of_votes)
  
  return(winner_results)
}


# Get the winner results for all wards
wardWinners <- tibble()
for (ward in 1:23) {
  wardWinner <- GetWinnerResults(ward)
  wardWinners<- bind_rows(wardWinners, wardWinner)
}
wardinfo <- read_csv("ward_info_2011.csv")
wardinfo <- wardinfo %>%
  select(area)

census <- read_csv("census_2016.csv")

wardFraction <- wardWinners %>%
  mutate(turnout = dataset$turnout) %>%
  mutate(population_density = census$Population/wardinfo$area)%>%
  mutate(household_density = census$Population/census$Households)

wardFraction <-wardFraction %>%
  mutate(area = cut(household_density, 
                               breaks = quantile(household_density,
                                                 probs = c(0, 1/3, 2/3, 1)),
                               labels = c('urban', 'suburban', 'rural'),
                               include.lowest = TRUE))
wardFraction %>%
  ggplot(aes(x = household_density,
             y = fraction_of_votes,
             color = area)) +
  geom_point(size = 3) +
  geom_smooth(se = FALSE, method = 'lm') + 
  facet_grid(rows = vars(area)) +
  labs(
    x = 'Average number of people per household',
    y = 'Fraction of votes winner got / Total amount of votes in a ward',
    color = 'Area',
    title = 'Average household vs fraction of votes have'
  )
```


In this plot, the density of people per household is put against the amount of people that actually voted. There are 3 graphs each representing the 3 different area types.
In the first plot, it is shown that as the household density increases, the fraction of votes won decreases. In the second graph and third graph representing higher household density, it is shown that as the household density increases, so does the fraction of votes won. This is the opposite conclusion from the first graph and the hypothesis. Hoever, the overall fraction of votes in the second graph is lower than the first and the third graph is the lowest. 

# Comments

 The fraction rate by was calculated by taking the amount of votes the winner had over the total amount of votes.  A low amount of fraction votes is from 0.35 to 0.50.
 A medium amount of fraction votes is from 0.50 to 0.65.  A high amount of fraction votes is above 0.65. The household density was split off into three sections to emphasize how as the house hold density increases between the 3 sections could affect the relationship. As shown in the graphs, it is not as simple as a straight line relationship. As shown by the first graph, the low amount of household showed the opposite conclusion when compared to the other two graphs. The first graph also carries the highest amount of data points and thus will have lower error and imporved stability.

Many precincts were removed from the data frame because they showed illogical data. Data points such as turnout rate being over 100% or infinit and fraction of votes won by being infinite. Their can not be more cards cast than registered voters.Outliers are included in the graph but are mostly seen when the total amount of votes is low.
This makes sense because low amount of data points gives rise to more variance. In the first graph, it is shows some data points with extremely high fraction of votes won by and turnout rate.

Results do not make sense in the first graph but the data shows the first graph is most likely incorrect. The results do makes sense in the second and third graph. We are mostly sure that as the fractions of votes won by decreases, the turnout increases. Varying results could happen due to population change, density between all the precinct. Assumptions include that all votes were accounted for fairly. All advance and special precints are removed from the dataset since the had 0 votes. Other assumptions include the precints are all anaylized fairly. 


# Suggestions for further explorations

-Ensure all presincts votes were counted properly and fairy will help the results be more accurate. In order to perhaps get a more accurate results. We could use a multivarable linear regression model. With more independing factors such as population density and numbers, and household density.


# Suggestions for further explorations

-Ensure all presincts votes were counted properly and fairy will help the results be more accurate.




# PART 2

#Comments

The winner rate by was calculated by taking the amount of votes the winner had over the total amount of votes. A high fraction of votes won by would be between 0.65 -0.8. A low fraction would be from 0.4-0.55. A medium amount of fraction of votes would be from 0.55 to 0.65 The precinct_winners were split off into three sections to emphasize how the household density could alter the data. A difference of 600 total votes between every precinct is a huge amount that could lead to a wrong conclusion. As shown by the first graph, the low amount of total votes showed the opposite conclusion when compared to the other two graphs. This is because an increase in the total amount of votes improves stability and lowers error. The household density ranges from 1.6-2.1 for urban. Sub-urban is 2.1-2.5. Rural is 2.5-2.9 roughly

Many precincts were removed from the data frame because they showed illogical data. Such as having 0 votes.
Results do make sense in the first graph but the data shows.. The results do  not makes sense in the second and third graph. We are mostly sure that as the household density increases, the fraction of votes decreases. Varying results could happen due to population change, density between all the wards. Assumptions include that all votes were accounted for fairly. All advance and special precints are removed from the dataset since the had 0 votes. Other assumptions include the precints are all anaylized fairly. 


# Suggestions for further explorations

-Ensure all presincts votes were counted properly and fairy will help the results be more accurate. In order to perhaps get a more accurate results. We could use a multivarable linear regression model. With more independing factors such as population density.


# References

No sources were used

# Appendix: sessionInfo
```{r}
sessionInfo()
```