---
title: "SDS 3386"
subtitle: "Explority analysis of the affect of greenhouse gas emissions"
author: "Shaafici Ali","Thinh Hong","Vinh 
date: "`r Sys.time()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---


```{r, warning=FALSE, message=FALSE}
# load libraries
library(dplyr)
library(janitor)
library(ggplot2)
library(skimr)
library(tidyverse)
library(broom)
library(readxl)
library(Rfast)
```
# Description
 By studying the data in the statementofvotescastoctober242018.xls,it is being deterrmined that if the winner and runner up are close in votes, there will be a higher turnout rate. The winner and runner up are compared in every precinct. Advance and special precincts are removed.
# Headline

The following plots is the turnout rate vs the fraction of the winner votes over the runner up votes. All the precinct winners were split into three sections by the total amount of votes. The first plot contains all precinct information with 0-(1/3) of the full range of total votes. Second plot is similar except it is split from 1/3 - 2/3.
Lastly the last plot is split from 2/3 to 3/3.
```{r, warning=FALSE, message=FALSE}

# Read dataset 
dataset <- read_excel("statementofvotescastoctober242018.xls", sheet = "Summary",skip = 2)

# Cleaning dataset and getting the turnout rate
dataset <- dataset[-c(1,2),]
dataset <- dataset %>%
  select(-(...3))%>%
  filter(!grepl('Adv',Precinct))%>%
  clean_names() %>%
  na.omit() %>%
  filter(precinct != "City / Ville - Total") %>%
  mutate(turnout = cards_cast / registered_voters) 

# Function to clean every ward with uneccesary rows. Also removes special precints
wardCleaner <- function(sheetNumber){
  # reads datasheet
  data <- read_excel("statementofvotescastoctober242018.xls", sheet = sheetNumber,skip = 2)
  #Cleans data
  data <- data %>% 
    clean_names() %>%
    select(-starts_with("x")) %>%
    select(-precinct_7,-overvotes,-undervotes,-times_cast) %>%
    filter(precinct_1 != "City / Ville") %>%
    filter(precinct_1 != "Ottawa") %>%
    filter(!grepl('Adv',precinct_1))%>%
    filter(!str_detect(precinct_1, "City / Ville - Total")) %>%
    rename(precinct = precinct_1)
  return (data)
}
# function to get winner votes and runner up votes in every presinct by ward
GetWinnerResults <- function(ward){
  sheet_data <- wardCleaner(str_c('Ward ', toString(ward)))
  #Creates empty tibble to add every precinct winner
  precint_winners <- tibble()
  n <- nrow(sheet_data)
  # for loop that goes through everyward per precinct
  for (i in 1:n){
  row_data <- sheet_data[i,]
  precint_winner <- row_data %>%
  #cleans data
  select(-precinct) %>%
  select(-total_votes) %>%
  # pivots to colums and renames
  pivot_longer(
      cols = everything(), 
      names_to = 'precinct',
      values_to = 'votes'
    )%>%
  # Adds votes for the winner and runner up at colums
 summarise(
   votes_winner = max(votes),
   votes_runner = Rfast::nth(votes, 2, descending = T)
 )
  #binds precinct winner to precinct winners
  precint_winners <- bind_rows(precint_winners, precint_winner)
  }
return (precint_winners)
}
#Creats emtpy tibble
precinct_winners <- tibble()
# Creates a for loop for every ward and binds the rows to the tibble
for (ward in 1:23) {
  precinct_winner <- GetWinnerResults(ward)
  precinct_winners<- bind_rows(precinct_winners, precinct_winner)
}

# Adds 3 new colums, fractions of votes the winner had more, turnout rate and total_votes. Then cleans data
precinct_winners <- precinct_winners %>%
  mutate(
    fraction_won_by = (votes_winner /votes_runner),
    turnout_rate = (dataset$turnout),
    total_votes = (dataset$cards_cast)
  )%>%
  subset(!turnout_rate == Inf)%>%
  subset(!fraction_won_by == Inf) %>%
  subset(!turnout_rate > 1)%>%
  na.omit()

# splits precinct_winners by the total votes into thrids
by_total_votes <-
  precinct_winners %>% 
  mutate(total_votes_type = cut(total_votes, 
                               breaks = quantile(precinct_winners$total_votes,
                                                 probs = c(0, 1/3, 2/3, 1)),
                               labels = c('low', 'middle', 'high'),
                               include.lowest = TRUE))
# plots all 3 total_votes_type
by_total_votes %>% 
  ggplot(aes(x = fraction_won_by,
             y = turnout_rate,
             color = total_votes_type)) +
  geom_point(size = 1) +
  geom_smooth(se = FALSE, method = 'lm') + 
   facet_grid(rows = vars(total_votes_type)) +
  labs(
    x = 'Fraction of votes winner had over runner up',
    y = 'Turnout Rate',
    color = 'total_votes_type',
    title = 'Relative amount of votes winner had over turnout rate'
  )

# creates a subset of precinct_winners by total vote type
low  <- subset(by_total_votes, total_votes_type =='low')
#cleans data
low  <- low %>%
  select(-votes_winner,-votes_runner,-total_votes_type)

#creats linear model with fraction won by as dependend and turnout rate as independent 
model_fit_low <- lm(fraction_won_by ~ turnout_rate, data = low)
# final summary of the wardWinners linear model
summary(model_fit_low)

middle <- subset(by_total_votes, total_votes_type =='middle')
middle  <- middle %>%
  select(-votes_winner,-votes_runner,-total_votes_type)

model_fit_middle <- lm(fraction_won_by ~ turnout_rate, data = middle)
summary(model_fit_middle)

higher <- subset(by_total_votes, total_votes_type =='high') 
higher  <- higher %>%
  select(-votes_winner,-votes_runner,-total_votes_type)

model_fit_higher <- lm(fraction_won_by ~ turnout_rate, data = higher)
summary(model_fit_higher)
```


In this plot, the turnout fraction is put against the fraction of which the winner beat the runner up. In the first graph representing the low section of total votes, it appears that as the fraction of votes won by increases, the turnout also increases. This is the opposite of the hypothesis. The second graph representing the middle of total votes shows a line that when the fraction of votes won by increases, the turnout decreases. The same as our hypothesis. Lastly, the third graph shows results similar to the second graph with lower variance.

# Comments

Precints were looked at instead of wards to increase amount of data. Larger data points allow for a more accurate result to the normal distrubution and draw a better conclusion. The winner rate by was calculated by taking the amount of votes the winner had over the amount of votes the runner up had. A high fraction of votes won by would be more than 3. A low fraction would approach 1 from a higher value. A high turnout rate would be above 0.5. A lot turnout rate is less than 0.25. A precinct is close if the fraction of votes won by is around 100%-120%. The precinct_winners were split off into three sections to emphasize how the total amount of votes could alter the data. A difference of 600 total votes between every precinct is a huge amount that could lead to a wrong conclusion. As shown by the first graph, the low amount of total votes showed the opposite conclusion when compared to the other two graphs. This is because an increase in the total amount of votes improves stability and lowers error.
In order to prove there is a correlation between turnout rate and by how much the winner beat the runner up the p- value and graph is looked at.

In the first graph representing the low section of total votes, it appears that as the fraction of votes won by increases, the turnout also increases. This is the opposite of the hypothesis. However, the data from the linear model of the low section shows a p-value of 0.00046 and a Adjusted R-squared of 0.067. Since the p-value is lower than 0.05, there is most likely not a difference when comparing turnout rate and fraction won by. The graph and p-value do not coincide. The data from the linear model of the middle gives a p-value of 0.07809. Proving there is most likely a correlation. The correlation being a close precinct increasing turnout rate.The second graph and p-value show the same thing. Lastly, the high total amount of votes has a p-value is 0.358. This shows there is even a more likely chance that as the fraction of votes won by decreases , the turnout increases. All graphs are compared to a 95% difference interval. The total amount of votes range roughly from 0-1800. So a low amount of votes is between 0-600, middle is 600,1200 and high is 1200-1800.

Many precincts were removed from the data frame because they showed illogical data. Data points such as turnout rate being over 100% or infinit and fraction of votes won by being infinite. Their can not be more cards cast than registered voters.Outliers are included in the graph but are mostly seen when the total amount of votes is low.
This makes sense because low amount of data points gives rise to more variance. In the first graph, it is shows some data points with extremely high fraction of votes won by and turnout rate.

Results do not make sense in the first graph but the data shows the first graph is most likely incorrect. The results do makes sense in the second and third graph. We are mostly sure that as the fractions of votes won by decreases, the turnout increases. Varying results could happen due to population change, density between all the precinct. Assumptions include that all votes were accounted for fairly. All advance and special precints are removed from the dataset since the had 0 votes. Other assumptions include the precints are all anaylized fairly. 


# Suggestions for further explorations

-Ensure all presincts votes were counted properly and fairy will help the results be more accurate. In order to perhaps get a more accurate results. We could use a multivarable linear regression model. With more independing factors such as population density and numbers, and household density.


# References

Learned from other students including Shaafici Ali and Quan vinh for winnerResults
# Appendix: sessionInfo
```{r}
sessionInfo()
```