---
title: "SDS 3386"
subtitle: "Assignment 3"
author: "Thinh Hong"
date: "`r Sys.time()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---


```{r, warning=FALSE, message=FALSE}
# load libraries
library(dplyr)
library(readxl)
library(janitor)
library(ggplot2)
library(skimr)
library(stringr)
library(plyr)
library(tidyverse)
```
# Description
 By studying the data in the statementofvotescastoctober242018.xls,it is being deterrmined  by what fraction of precints did the winner beat the 2nd place person in all the wards. The amount of precints the winner beat the runner up against the total amount of precints in each ward will be ploted. There are 23 wards.

# Headline

The following plot is  the comparision between the first place winner and the second place winner. It shows the percentage of precincts that the first place winner beat the runner up

```{r, warning=FALSE, message=FALSE}

dataset <- function(sheetNumber){
  sheet <- read_excel("statementofvotescastoctober242018.xls", sheet = sheetNumber, skip = 2)
  sheet <- sheet %>% clean_names %>% filter(precinct_1 != 'City / Ville') %>% filter(precinct_1 != "Ottawa") %>% select(-"precinct_7", -"times_cast",-"undervotes",-"overvotes", - "total_votes") %>% select(-starts_with('x'))
  return (sheet)
}

determineWinner <- function(sheet){
  
  wardData <- dataset(sheet)
}

```


```{r}

data <- read_excel("statementofvotescastoctober242018.xls", sheet = 3, skip = 2)
data <- data %>% clean_names %>% filter(precinct_1 != 'City / Ville') %>% filter(precinct_1 != "Ottawa") %>% select(-"precinct_7", -"times_cast",-"undervotes",-"overvotes") %>% select(-starts_with('x'))

sheet <- tail(data,1)

```
In this plot, the fraction of wins of precincts is showed in a bar graph for all 23 wards. It is being determined if the winner of the ward beat the runner up of the ward in how many precincts against the total precincts. Here we can see that the percentage of precincts the winner wins varies a lot between wards. From 40% to 80% wins.

# Comments
In order to be consider winning in a precinct, the total amount of votes for every candidate is calculated and the highest amount wins. 
The results do makes sense as the winner overall wins over 50% of the precints. Reasons that could be true when the percentage wins was below 50 is the sample size could be very small. Either in population or votes. A smaller sample size would make the results vary more than other wards.


# Suggestions for further explorations

-Ensure all presincts votes were counted properly and fairy will help the results be more accurate.




# PART 2

#Description
We are using a proportional test to see if this coin toss is fair. 

# Headline
The following proportional test is comparing all the coin tosses and seeing if the coin toss is fair. The results of the coin toss are given. The null hypothesis is assuming that the coin toss is fair and there is no true difference in the means between head and tails. The alternate hypothesis states that the coin toss is not fair and there is a true difference in means between heads and tails. 

```{r}
x <- c("t", "t", "h", "t", "h", "h", "t", "t", "t", "t", "t", "t", "h", "h", "t", "h", "t", "t", "t", "t", "t", "h", "t", "t", "h", "t")

hs = str_count(x,"h")
ts = str_count(x,"t")
heads <- sum(hs,1)
tails <- sum(ts,1)

prop.test(8,26, p = NULL, alternative = "two.sided", correct = TRUE)
prop.test(18,26, p = NULL, alternative = "two.sided", correct = TRUE)
```

#Comments

We are assuming that the probability of the coin landing on heads is 50%. We are also assuming the coin is being tossed fairly. The coin being heads shows that it was a success. According to the binomial test, the p value is 0.07552. Which is also the same p- value that was gotten from the previous test(binomial). Which indicates that the coin toss is indeed fair. We fail to reject the null hypothesis. If the null hypothesis should be rejected but was not, A type II error will have occured. We are 95%  sure there is no true difference. We are also assuming that the coin was tossed 26 times with no set goal. The coin was not tossed untill 8 heads or 18 tails were made. 

# Suggestions for further explorations

In order to have a higher chance of correctly assuming the coin toss is fair, doing more trials will have bring the probability to where it should be.

# References

No sources were used

# Appendix: sessionInfo
```{r}
sessionInfo()
```